---
title: "Compare to Anne's Previous Results"
output: 
  html_notebook:
    toc: true
  html_document:
    toc: true
    toc_float: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
start_time <- Sys.time()
```

Anne found quite a large number of discrepancies between her previous results
and what she got from the HatcheryPedAgree workflow.  I am going to try to understand
those differences here.

I ran my own HatcheryPedAgree workflow in `typical-analysis-RR.Rmd`, and I stored
various outputs in an Unsynced directory.
```
(base) /HatcheryPedAgreeOutputs/--% ls -R RussianRiver-July-6-2020/
Anne_prior_results/ for_read.rds        reorg.rds           snppit-run-dir/

RussianRiver-July-6-2020//Anne_prior_results:
RR_parentage_corrected-sexes_June27_2019.csv
Sad.rds
noSad.rds

RussianRiver-July-6-2020//snppit-run-dir:
BigSmax_Input                           snppit_output_ChosenSMAXes.txt
BigSmax_Output.txt                      snppit_output_FDR_Summary.txt
PurePop_Input                           snppit_output_ParentageAssignments.txt
PurePop_Output.txt                      snppit_output_PopSizesAnPiVectors.txt
snppit_input.txt                        snppit_output_TrioPosteriors.txt
snppit_output_BasicDataSummary.txt      snppit_seeds

```
That includes `Anne_prior_results`, which has Anne's previous results,
and also `snppit-run-dir`, which has all the snppit output from my
HatcheryPedAgree run, and also the `for_read.rds` (which is the `for_real`
variable, and `reorg.rds` which is the `reorg` variable).

# Libraries

```{r}
library(tidyverse)
library(HatcheryPedAgree)
```

# Get anne's previous results

```{r}
apr <- read_csv(
  "~/Documents/UnsyncedData/HatcheryPedAgreeOutputs/RussianRiver-July-6-2020/Anne_prior_results/RR_parentage_corrected-sexes_June27_2019.csv", 
  guess_max = 20000)
```

# Get the noSAD results that I computed

Might as well go straight to the noSAD results...

```{r}
sex_and_date <- read_rds("private/rrsh_metadata.rds")
my_nosad <- slurp_snppit(
  DIR = "~/Documents/UnsyncedData/HatcheryPedAgreeOutputs/RussianRiver-July-6-2020/snppit-run-noSAD",
  S = sex_and_date
)
reorg <- read_rds("~/Documents/UnsyncedData/HatcheryPedAgreeOutputs/RussianRiver-July-6-2020/reorg.rds")
```

Now, the first thing I want to investigate are those instances where Anne found a parent
pair and the HPA workflow did not.  We have the wrinkle that we use only a single name for
each individual, even if they were genotyped multiple times (as a "matching" sample).

In other words, the individual might be called something different in Anne's data set
than in the HPA data set.  So, to get around this, I want to augment the no_sad results
I got, by including rows in it that represent all the different ways a trio
(kid, pa, and ma) could have been named (that are different from the canonical way
that HPA used), so that we can left join them (as kids), and also properly find them
as parents.

So, for each canonical name, we need to get all the other, **different** names that
might be used in Anne's results.  At first, we just deal with getting all the extra names
for the kids.  We deal with ma and pa later.
```{r}
aliases <- reorg$matchers_metadata %>%
  filter(new_id != original_id) %>%
  select(original_id, new_id)
kid_additions <- aliases %>%
  left_join(my_nosad, by = c("new_id" = "kid")) %>%
  rename(kid = original_id) %>%
  select(-new_id)

my_nosad2 <- bind_rows(my_nosad, kid_additions)
```


# join anne's results

For now, I just want to focus on what anne found, so we left join:
```{r}
aprLJ <- left_join(
  apr, 
  my_nosad2, 
  by = c("Kid_ID" = "kid"),
  suffix = c("_apr", "_hpa")
  )
```

Now, quickly look at the FDRs here:
```{r}
ggplot(aprLJ, aes(x = FDR_apr, y = FDR_hpa)) +
  geom_point(alpha = 0.1) +
  theme_bw()
```


```{r}
aprLJ %>%
  filter(Kid_ID %in% aliases$original_id ) %>%
  View()
```


```{r}
looksie <- aprLJ %>%
  select(Kid_ID, Ma_ID, ma, Pa_ID, pa, FDR_apr, FDR_hpa, Pvalue_apr, 
         Pvalue_hpa, Kid_Year, Ma_SpawnYear, Kid_Hatchery, MaxP.Pr.Relat_apr, MaxP.Pr.Relat_hpa) %>%
  mutate(correct = (ma == Ma_ID & pa == Pa_ID) | (pa == Ma_ID & ma == Pa_ID) )
```

```{r}
ggplot(looksie %>% filter(correct == TRUE), aes(x = Pvalue_apr, y = Pvalue_hpa, colour = MaxP.Pr.Relat_hpa)) +
  geom_point(alpha = 0.5) + 
  theme_bw() +
  geom_abline(slope = 1, intercept = 0) +
  facet_grid(Kid_Hatchery ~ Ma_SpawnYear)# +
  #ylim(0, 0.15)
```

OK,  that is interesting.  I think it might have to do with using pooled allele frequencies.
Let's look at those allele freqs. We can do that here, in R.

## Allele freqs

Here we can look at the allele frequencies that SNPPIT is using to compute likelihoods.
I think HPA uses all the fish except for those in the last year.
```{r}
genos <- read_rds("private/rrsh_genotypes.rds")
meta <- read_rds("private/rrsh_metadata.rds")

gmeto <- genos %>%
  left_join(meta, by = "indiv")

grand_mean <- gmeto %>%
  filter(!is.na(allele_int)) %>%
  count(locus, allele_int) %>%
  group_by(locus) %>%
  mutate(gm_freq = n / sum(n))

hatchery_means <- gmeto %>%
  filter(!is.na(allele_int)) %>%
  group_by(hatchery, locus, allele_int) %>%
  tally() %>%
  mutate(freq = n / sum(n)) %>%
  left_join(grand_mean, by = c("locus", "allele_int"))
```
Let's plot those and see what it looks like:
```{r}
ggplot(hatchery_means, aes(x = gm_freq, y = freq)) +
  geom_point() + 
  geom_abline(intercept = 0, slope = 1) +
  facet_wrap(~ hatchery, nrow = 1)
```

OK, now what I would like to do is see how things compare over kid return years from 2007 to 2016.  
We will make a function for that.  I want to compare the frequencies that Anne used (aggregated over
hatcheries, but only including the parents from 1 to 7 years previous) to the means (over all years)
for each of the hatcheries.  And I also want to get the actual frequencies in each of the hatcheries
with just those 7 years included each time.

```{r}
agg_freqs_by_kidyear <- function(D, kyear) {
  D %>%
    filter(year >= kyear - 7, year <= kyear - 1) %>%
    filter(!is.na(allele_int)) %>%
    group_by(locus, allele_int) %>%
    tally() %>%
    mutate(
      agg_freq = n / sum(n),
      kidyear =  kyear
      )
}
hatch_freqs_by_kidyear <- function(D, kyear) {
  D %>%
    filter(year >= kyear - 7, year <= kyear - 1) %>%
    filter(!is.na(allele_int)) %>%
    group_by(hatchery, locus, allele_int) %>%
    tally() %>%
    mutate(
      hatch_freq = n / sum(n),
      kidyear =  kyear
      )
}

hatch_freqs <- lapply(2007:2016, function(x) hatch_freqs_by_kidyear(gmeto, x) ) %>%
  bind_rows()
agg_freqs <- lapply(2007:2016, function(x) agg_freqs_by_kidyear(gmeto, x) ) %>%
  bind_rows()
```

Now, we can plot some of these things.  First, let's plot the aggregated freqs Anne used
against the overall hatchery freqs:
```{r}
agg_freqs %>%
  left_join(hatchery_means, by = c("locus", "allele_int")) %>%
  ggplot(aes(x = freq, y = agg_freq)) +
  geom_point() +
  facet_grid(kidyear ~ hatchery)
```

Nothing super revealing there.  It does show that the variability in Coyote Valley is a little higher
and that might account for what look like more FDR discrepancies between HPA and the pooled allele freqs
approach in Coyote Valley (at least, looking at the the figures a few sections up.)
