---
title: "Developing Sex Inference"
output: html_notebook
---

# Introduction

I want to quickly see what the landscape looks like for this.
Basically, we need to have connected components where sex
is known for at least one parent.

I will just work from some RR data at first:
```{r, message=FALSE}
library(tidyverse)
library(tidygraph)
library(HatcheryPedAgree)
library(ggraph)

rr <- read_rds("../private/rr-results/nosad.rds") %>%
  filter(FDR < 0.01, MaxP.Pr.Relat == "C_Se_Se")

# note that if count up how many sex mismatches there are
# you see there are a lot!
rr %>% 
  filter(ma_sex == pa_sex | ma_sex == "Male" | pa_sex == "Female") %>%
  nrow(.)
```

But, most of those are just because the ma and pa columns have not
been organized as they should:
```{r}
rr_fix <- reformat_no_sex_or_date_results(rr)

# Now, all the problematic cases are where pa and ma are the same sex
rr_fix %>% 
  filter(ma_sex == pa_sex | ma_sex == "Male" | pa_sex == "Female")

```
Not only that, but we do have some individuals with unknown sex:
```{r}
rr_fix %>%
  filter(is.na(pa_sex) | is.na(ma_sex))
```
Note also that some indivdiuals are named with a trailing `_HathceryName` that
we will want to remove to start working with them.

Our goal is to use the existing known matings to do our best to try
to correct sexing errors.  For example, if someone who has been called a male
is mated to three different individuals that are males, then it is probably
best to recode that male as a female.  

We can end up coding all this up in terms of the Sum Product algorithm on
a factor graph that connects mating pairs (to find the marginal probability)
of each individual's sex.  

That should be easy to code up because each factor node (a marriage node) will 
have exactly two variable nodes (the two mates) attached to it.  There is
no need to put offspring into this.

What's more, we will have mostly small connected components, and we probably
won't have too many cycles.  However, I want to explore that now.  

In the end, I think that just looking at it and doing it by eyeball might
work better.


# Investigating graphical structure / connectedness, etc

First, we want to give a canonical name for each pair of mates.
We do this by removing the `_Hatchery` on names where it exists and
then sorting the names lexicographically
```{r}
cano_rr <- rr_fix %>%
  mutate(
    tmp1 = str_replace(pa, "_.*$", ""),
    tmp2 = str_replace(ma, "_.*$", ""),
  ) %>%
  mutate(
    cano1 = case_when(
      tmp1 < tmp2 ~ tmp1,
      TRUE ~ tmp2
    ),
    cano2 = case_when(
      tmp1 < tmp2 ~ tmp2,
      TRUE ~ tmp1
    ),
    sex1 = case_when(
      tmp1 < tmp2 ~ pa_sex,
      TRUE ~ ma_sex
    ),
    sex2 = case_when(
      tmp1 < tmp2 ~ ma_sex,
      TRUE ~ pa_sex
    )
  ) %>%
  select(cano1, cano2, sex1, sex2, everything())
```

Now we can enumerate the pairs:
```{r}
pairs <- cano_rr %>%
  count(cano1, cano2)
```

And now we want to make a graph out of those mated pairs, and determine
the connected component that each individual is in. To find the connected components
we don't even need to put the marriage nodes in there...
```{r}
conn_comps <- pairs %>%
  rename(from = cano1, to = cano2) %>%
  select(to, from) %>%
  igraph::graph_from_data_frame() %>%
  igraph::components() %>%
  .$membership %>%
  enframe() %>%
  rename(indiv = name, cluster = value) %>%
  arrange(cluster, indiv) %>%
  group_by(cluster) %>%
  mutate(cluster_size = n()) %>%
  ungroup()
  
```
So, let's look at the size of the different conn_comps:
```{r}
conn_comps  %>%
  group_by(cluster) %>%
  summarise(n = cluster_size[1]) %>%
  ggplot(aes(x = n)) + 
  geom_histogram(binwidth = 1)
```

So, that looks about right.  Except what is up with connected component of size 65 or so?
When and where does that occur, and why?  Let's make some plots of all of those.  To do that
we will make a tidy graph of the mates, and then attach some information to
the nodes.
```{r}
tgraph <- pairs %>%
  rename(from = cano1, to = cano2) %>%
  select(to, from) %>%
  igraph::graph_from_data_frame() %>%
  as_tbl_graph()

# get sex info for each fish from the info in the pedigree
sex_info <- bind_rows(
  rr_fix %>% select(kid, kid_sex) %>% rename(indiv = kid, sex = kid_sex),
  rr_fix %>% select(ma, ma_sex)  %>% rename(indiv = ma, sex = ma_sex),
  rr_fix %>% select(pa, pa_sex) %>% rename(indiv = pa, sex = pa_sex)
) %>%
  count(indiv, sex) %>%
  select(-n)

# check to make sure that no individual has multiple sexes
sex_info %>%
  count(indiv) %>%
  filter(n > 1)
  
```
Now, also from the pedigree, let's get a data frame that has
one row for every year that an individual was spawned.
```{r}
spawn_info <- bind_rows(
  rr_fix %>% select(ma, ma_year)  %>% rename(indiv = ma, spawn_year = ma_year),
  rr_fix %>% select(pa, pa_year) %>% rename(indiv = pa, spawn_year = pa_year)
) %>%
  count(indiv, spawn_year) %>%
  select(-n)

# oddly, this doesn't seem to have any fish that spawned in
# multiple years.  WTF?
spawn_info %>%
  count(indiv) %>%
  filter(n > 1)

```
Now, join that info onto the nodes of the graph:
```{r}
tgraph2 <- tgraph %>%
  left_join(sex_info, by = c("name" = "indiv")) %>%
  left_join(spawn_info, by = c("name" = "indiv")) %>%
  left_join(conn_comps, by  = c("name" = "indiv"))
```

That leaves us with 11 clusters.  Let's see if we can plot them
in a faceted graph.
```{r}
tgraph2 %>%
  filter(cluster_size > 15) %>%
  ggraph(layout = "kk") +
  geom_node_point(aes(shape = sex, fill = as.factor(spawn_year))) + 
  geom_edge_link() +
  scale_shape_manual(values = c(Male = 22, Female = 21), na.value = 23) +
  facet_nodes(~ cluster, scales = "free")
```

# Clusters with clashing-sex pairs

That is pretty cool.  So, what does it look like if we investigate all
the clusters in which there are individuals that were of clashing
sex:
```{r}
tmp <- rr_fix %>% 
  filter(ma_sex == pa_sex | ma_sex == "Male" | pa_sex == "Female")
sex_pair_clashers <- tibble(
  indiv = unique(c(tmp$pa, tmp$ma)),
  sex_clasher = TRUE
)

  

tgraph3 <- tgraph2 %>%
  left_join(sex_pair_clashers, by = c("name" = "indiv")) %>%
  mutate(sex_clasher = ifelse(is.na(sex_clasher), FALSE, sex_clasher))

sex_clash_clusters <- tgraph3 %>%
  group_by(cluster) %>%
  filter(any(sex_clasher == TRUE)) %>%
  ungroup()

  
```
Then plot those:
```{r}
g <- ggraph(sex_clash_clusters, layout = "kk") +
  geom_edge_link() +
  geom_node_point(aes(shape = sex, fill = sex_clasher, colour = sex_clasher), size = 10) + 
  geom_node_point(aes(shape = sex), fill = NA, colour = "black", size = 10) + 
  geom_node_text(aes(label = name), repel = TRUE) +
  facet_nodes(~ cluster, scales = "free") +
  scale_shape_manual(values = c(Male = 22, Female = 21), na.value = 23)

ggsave(g, filename = "outputs/sex-clasher-facet-graphs.pdf", width = 40, height = 40)
```
That is cool.  There are some that clearly can be resolved by changing
the sex of a single individual.


# Cluster with unknown sex pairs

We should be able to do the same with the clusters that include
fish that don't have a reported sex.
```{r}
sex_unknown_clusters <- tgraph3 %>%
  group_by(cluster) %>%
  filter(any(is.na(sex))) %>%
  ungroup()
```
Then plot:
```{r}
g <- ggraph(sex_unknown_clusters, layout = "kk") +
  geom_edge_link() +
  geom_node_point(aes(shape = sex, fill = sex, colour = sex_clasher), size = 10) + 
  geom_node_point(aes(shape = sex), fill = NA, colour = "black", size = 10) + 
  geom_node_text(aes(label = name), repel = TRUE) +
  facet_nodes(~ cluster, scales = "free") +
  scale_shape_manual(values = c(Male = 22, Female = 21), na.value = 23) +
  scale_fill_manual(values = c(Male = "skyblue", Female = "pink"), na.value = "white")

ggsave(g, filename = "outputs/sex-unknown-facet-graphs.pdf", width = 40, height = 40)
```
