---
title: "Typical Analysis and Testing...RR"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: paged
  html_notebook:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
start_time <- Sys.time()
```


## Load the Data

```{r}
library(tidyverse)
library(HatcheryPedAgree)

genotypes <- read_rds("../private/rrsh_genotypes.rds")
metadata <- read_rds("../private/rrsh_metadata.rds")

```

## Assess Patterns of Missing Data, and deal with it

```{r}
miss_dsn <- genotypes %>%
  group_by(indiv) %>%
  summarise(
    num_all_loci = n() / 2,  # check to make sure each gene copy is explicitly listed as NA
    num_non_miss_loci = sum(!is.na(allele_int)) / 2)

count(miss_dsn, num_all_loci)  # good.  Everyone has 94 loci (even if some are missing)

full_histo <- ggplot(miss_dsn, aes(x = num_non_miss_loci)) +
  geom_histogram(binwidth = 1)
full_histo
```
So, these data are already filtered to have >= 85 non-missing loci.

## Matching Samples

```{r}
meta <- metadata
genos <- genotypes

# first time through, just get the distribution of matching proportions
for_histo <- find_matching_samples(genos, min_frac_matching = 0.8)

# check the distribution
for_histo$pairs %>%
  mutate(frac_match = num_match / num_non_miss) %>%
  ggplot(aes(x = frac_match)) +
  geom_histogram()

# get the clusters of matching genotypes each indvidual belongs to
for_real <- find_matching_samples(genos, min_frac_matching = 0.95, return_clusters = TRUE)

# we will end up using the identified clusters
head(for_real$clusters)
```

I will save that for later
```{r}
write_rds(for_real, path = "~/Documents/UnsyncedData/HatcheryPedAgreeOutputs/RussianRiver-July-6-2020/for_read.rds")
```


## Reorganizing for SNPPIT

At this point, `for_real$clusters` is the tibble we need to re-organize our genotypes and meta-data
for SNPPIT.  For every cluster of matching samples we will use, as the genotype, the
sample with the least missing data.  We will also use the sex of that individual (sometimes
there are mismatches in the sex of the matching genotypes).  Sometimes there are mismatches in
the hatchery of the matching genotypes.  In those cases, each separate hatchery gets its own
canonical individual named as the ID of the main canonical individual with the hatchery
name appended to it.  The following function takes care of this and reorganizes both the
genotypes and also the meta data into `snppit_genos` and `snppit_meta`
(as well as a few other list components).
```{r}
reorg <- reorganize_matching_samples(
  genotypes = genos, 
  metadata = meta, 
  clusters = for_real$clusters
)

# and I will also save that
write_rds(
  x = reorg,
  path = "~/Documents/UnsyncedData/HatcheryPedAgreeOutputs/RussianRiver-July-6-2020/reorg.rds",
  compress = "xz")
```

Let's have a look at some of the different components of that output.

### `matchers_metadata`

This is the meta data for all the matching genotypes.  Column `original_id` shows what they
were named on input, and column `new_id` shows the ID used to identify them now in the
SNPPIT-ready output.
```{r}
reorg$matchers_metadata
```

### `snppit_meta` and `snppit_genos`

These are the tibbles that are ready to pass into `prepare_snppit_infile()`.  Multiple
years and spawner_groups of the matching individuals have been lumped into
comma-separated strings for the year and spawner group inputs to SNPPIT.

### `cross_hatchery_matches` 

A tibble that shows you which clusters of matching genotypes included fish from more
than one hatchery.
```{r}
reorg$cross_hatchery_matches
```

### `cross_sex_matches` 

A tibble that shows you which clusters of matching genotypes included fish with more than
one reported sex
```{r}
reorg$cross_sex_matches
```


## Prepare a SNPPIT infile and run it

We have rolled these two steps into a single run_snppit() function.

Internally, it calls `prepare_snppit_infile()` to write the data,
and then it runs `snppit` inside the `system()` command.

Here is what that looks like:
```{r}
snppit_dir1 <- run_snppit(
  reorg$snppit_genos, 
  reorg$snppit_meta, 
  outdir = "~/Documents/UnsyncedData/HatcheryPedAgreeOutputs/RussianRiver-July-6-2020/snppit-run-dir"
)

snppit_dir1
```


# Now, do an unconstrained run

We won't use sex or date (but year and hatchery still need to match).  And we
will set the possible age of maturity from 1 to 7.
```{r}
snppit_dir2 <- run_snppit(
  reorg$snppit_genos, 
  reorg$snppit_meta, 
  outdir = "~/Documents/UnsyncedData/HatcheryPedAgreeOutputs/RussianRiver-July-6-2020/snppit-run-noSAD",
  use_spawner_group = FALSE,
  use_sex = FALSE,
  min_age = 1,
  max_age = 7,
)

snppit_dir2
```

# Running Time

Running the code and rendering this notebook required approximately this much time
on a Mac laptop of middling speed:
```{r}
Sys.time() - start_time
```


