---
title: "Using the partition_two_snppit_runs() function"
author: "Eric C. Anderson"
output:
  html_document:
    df_print: paged
---

# Introduction

This is primarily used for identifying the different combinations
of results that you might see when comparing SAD (with Sex And Date)
to noSAD (without Sex and Date) runs.  Typically you will pass the results of the SAD
run in as Run1 to the function and the results of the noSAD run in as
Run2.  For each run you can set a different FDR threshold to use when
the function is partioning things, but I suspect that it will be typical
to set FDR1 = 0.01 and FDR2 = 0.01.  


This function simply explicitly partitions each kid from the pedigree inference
project into a group based on characteristics of the inference of its parents in the
two different snppit runs.  I find this easier to look at and think about than
trees depicting filtering decision points that may or may not represent actual partitions.

Briefly, the partition of these individuals is based on just a few binary characteristics
of each kid within each run.  Namely:

1. Does the kid have any inferred parents, or none (i.e. NA) in Run1 and in Run2?  If
it doesn't have any inferred parents in a run, then the later criteria (2, 3, and 4 are
largely moot).

2. Did the trios in Run1 and Run2 enjoy having their maximum posterior trio category be C_Se_Se, or not?

3. Was the trio found at an FDR below a certain threshold for each run?

And finally, a fourth criterion that is not binary like the first three, and, further, is
also a property of the comparison between the trios found for a kid in the two different runs,
rather than being a property that can be associated with just a single run:

4. How many parents are shared for the kid between the two runs, with possible answers 0 = (none!), 1 = just a single parent is shared, 2 = the same two parents are inferred in the two runs, or NA = in one or both of the runs the kid was not found to have a parent.

The total number of possible categories from criteria 1--3 that a kid in a single run can
belong to is: 1 (for NA) + 1 * 2 * 2.  So, at the first 3 criteria between the two 
Runs there are 25 different categories possible.  

Of those, 16 involve cases where parents are NA in neither run.  And in those cases there
are three possible numbers of matching parents (0, 1, or 2), which turns those 16 into a possible
48 cases.  

Hence, we are partitioning the kids/trios up into 48 + 9 = 57 possible categories with these few
criteria.

Once you get the number in each of those categories, seeing them all out there in a way
that adds up to the actual number of kids makes it a lot easier to think about which trios you want
to retain and which you don't.  In fact, as we will see, we can just add a column which says, for
each partition, whether you want to retain the trio from Run1, from Run2, or from Neither, then
feed that back to pick out the trios you want.  (I haven't implemented that yet, but it will be easy).

# Example using Laura's SAD and noSAD results

I got these from laura and have put them in `private`.  So it looks like this:
```{r, message=FALSE}
library(tidyverse)
library(HatcheryPedAgree)

SAD <- read_rds("../private/cv-results/CV_SAD_results.rds")
noSAD <- read_rds("../private/cv-results/CV_no_SAD_results.rds")

PR <- partition_two_snppit_results(SAD, noSAD)
```

Here you can see a few different lines of the resulting data frame of 19,028 kids, namely
one row from each of hte 29 different non-empty categories.  The
two new columns added are `bit_category` and `verbal_category`.

You have to click to the right to see all the columns.
```{r}
PR %>%
  group_by(bit_category) %>%
  slice_sample(n = 1)
```

Now, count up occurrences in the different partitions.  We sort them by number of occurrences:
```{r}
Summary <- PR %>%
  count(bit_category, verbal_category) %>%
  select(n, everything()) %>%
  arrange(desc(n))

Summary
```


Finally, it can be nice to break up the verbal description of the categories
into three columns (this allows easier sorting, etc.)  You can break them
up on the pipe: `|`, like so:
```{r}
Summ_sepa <- Summary %>%
  separate(
    col = verbal_category,
    into = c("verbal_bit", "run_1_stuff", "run2_stuff", "run1_vs_run2_num_shared_parents"),
    sep = "\\|"
  )
```

I write that as a CSV and send it to Anne, Laura, Carlos, and Devon.

```{r}
write_csv(
  Summ_sepa,
  path = "../development/outputs/cv-sad-vs-nosad-partitions.csv"
)
```

# Implementation Notes

57 might seem like a lot of categories to keep track of.  I relied heavily on
bit-masking to make it easier.  I probably could have gotten this code cleaner, but
here is all the new code for the function.  
```r
`r paste(readLines("../R/partition_two_snppit_results.R"), collapse = "\n")`
```

If anyone wants to delve into this further with me I can tell you how bitmasks work.

